Approximating Information-Theoretic Scores\\for Computer Adaptive Tests\\in Bayesian and Credal Networks

### 2021-05-06 paper.tex

>>>>

\section{A Toy Example}
As a first, demonstrative, example of our ideas consider a model characterized by a single (Boolean) Skill $S$ and a set of questions $\bm{Q}$ picked from a set of templates. Let us first focus on a Bayesian, i.e., precise, case. The prior specification $P(S)$ is achieve by a single parameter $p \in [0,1]$ such that $P(S=1)=p$. For a generic question $Q$, which is also assumed to be Boolean, a complete specification of the CPT $P(Q|S)$ is obtained by the two parameters $p_0,p_1\in [0,1]$ where $P(Q=1|S=0)=p_0$ and $P(Q=1|S=1)=p_1$. In other wrods $p_0$ denote the probability of providing a right answer to the question for a student who does not have the considered skill and $p_1$ for a student who have that skill. An obvious constraint is $p_1>p_0$, i.e., having the skill makes your probability of providing a right answer higher. The inequality should be strict, otherwise the two probabilities are equal, and this means that the answer is irrelevant for the skill. Although parameters $p_1$ and $p_0$ already give a complete parametrization, we perform the linear transformation
$\lambda:=1-\frac{1}{2}(p_0+p_1)$ and $\delta=(p_1-p_0)$. The first quantity is the arithmetic average of giving a wrong answer and we might indent it as the \emph{level} of the question, this mining that the higher is $\lambda$ the more unlikely is to receive a correct answer is the students have the same probability of having and not having the skill. Regarding $\delta$ we might intend it as the (average) discriminative power of a question.
Note that the inverse relations give $p_1=1-\lambda+\frac{\delta}{2}$
and $p_0=1-\lambda-\frac{\delta}{2}$. Where constraint $p_1>p_0$ is just $\delta>0$ and any other value in $[0,1]$ for both parameters is valid.

In our toy example, we consider nine possible templates, made of triads of templates with the same $\delta$ and same $\lambda$. The reference values $[0.4,0.5,0.6]$ are used.
